Transition path sampling based calculations of free energies of enzymatic reactions: the case of human methionine adenosyl transferase and plasmodium vivax adenosine deaminase

### 2022-05-10 tex/manuscript.tex
added some more stuff

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  Transition path sampling (TPS) has been widely applied for the
  calculations of reaction rate, transition state and reaction 
  coordinates of condensed phase systems. Here, we discuss a scheme 
  for the calculation of free energies with the TPS method in combination 
  with a window based sampling technique. This 
  distribution can then be used for the calculations of free energies
  using standard Boltzmann inversion schemes. We calculate the free
  energy profiles of the reactions catalyzed by the human methionine 
  S-adenosyltransferase (MAT2A) enzyme and the plasmodium 
  vivax adenosine deaminase enzyme to demonstrate the applicability of this 
  method. 
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Molecular dynamics (MD) simulations are increasingly being
used to calculate quantitative estimations of experimentally measurable 
kinetic and thermodynamic parameters of enzyme catalyzed reactions. 
\cite{Karplus02NatStructMolBiol9p646,
Schramm98AnnuRevBiochem67p693,Zhang05AccChemRes38p379,
Schramm11AnnuRevBiochem80p703,Schramm18ChemRev118p11194} 
This has helped in proposing mechanisms and explanations for 
enzyme activity, fast high throughput screening of thousands of inhibitor 
molecues in drug lead discovery, \cite{Jorgensen09AccChemRes42p724,Sliwoski14PharmacolRev66p334} 
directed evolution for designing enzymes with enhanced catalytic 
activities, \cite{Thyme09Nature461p1300,Bloom09PNAS106p9995,Schafer19JAmChemSoc141p10431} etc.    
Straightforward MD simulations adequately sample 
the long-lived stable reactant and product states 
but the high energy barrier-crossing events that occur rarely and at long 
time scales are not easily accessible using this method alone.
To access such states at a higher frequency using MD simulations 
several enhanced sampling techniques 
have been developed in the past few decades some of which are 
umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
steered MD, \cite{Park04JChemPhys120p5946}
milestoning, \cite{Faradjian04JChemPhys120p10880} and transition path sampling 
(TPS). \cite{Pratt86JChemPhys85p5045,Bolhuis02AnnRevPhysChem53p291,
Dellago98JChemPhys108p1964,Bolhuis21AdvTheory4p2000237}

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{figures/mat2a-reaction.png}
\caption{The reaction catalyzed by human MAT2A enzyme}
\label{fig:mat2a-reaction}
\end{figure}

$pv$ADA catalyzes the reaction between adenosine and hydroxide ion 
resulting in the formation of inosine and ammonia as shown in
Fig. \ref{fig:ada-reaction}.
$pv$ADA is an essential enzyme in the purine salvage pathway and 
inhibiting this enzyme is a promising approach to treat malaria.   
This reaction proceeds with the nucleophilic O atom (O$^{\text{hydroxyl}}$) of the hydroxide anion
attacking the electrophilic C6 atom of the adenosyl ring leading to the formation of a Meisenheimer 
complex while the N1 atom on the adenosyl ring starts to get 
protonated by the Glu229 residue of $pv$ADA. At the transition state, experimental findings 
suggest that the dissociation of the C6$-$N6 bond as well as the protonation of the N1 
atom are almost complete. \cite{Luo07JAmChemSoc129p8008} The key bond parameters are 
the distance between the approaching hydroxyl oxygen and the C6 atom (denoted as $dis(OC)$) and the 
dissociating C6$-$N6 ($dis(NC)$) bond distance. For the TPS simulaions we define the reactant state to 
have $dis(NC)$ being less than $1.7$ {\AA} and the product state to have the 
$dis(OC)$ less than $1.7$ {\AA}.

\begin{figure}
\includegraphics[scale=0.6]{figures/ada-new.png}
\caption{The reaction catalyzed by adenosine deaminase enzyme}
\label{fig:ada-reaction}
\end{figure}

A biased initial reactive trajectory is obtained by using harmonic restraint forces 
of 1500 to break the $O-C$ bond and form the $S-C$ bond for the MAT2A enzyme and 
to break the $N-C$ bond and form the $O-C$ bond for the case of the $pv$ADA enzyme. 
Using the biased trajectory as the starting point and after removing the restraint 
forces, the shooting algorithm was used to generate
new trajectories. The well known shooting algorithm begins with the choice of a 
random time slice from the biased reactive trajectory and perturbing the momenta of the system.
Then the dynamics is propagated for 250 steps forawards and backwards using $1\;fs$ time steps
resulting in a trajectory with a time period of $0.5\;ps$.
A total of 150 and 130 reactive trajectories were generated for the MAT2A and $pv$ADA
enzymes, respectively. 
%------------------------------------------------------------------------------ 
\subsection{Free energy calculations}
%------------------------------------------------------------------------------ 
To obtain the free energy profile along the order parameter, we use the method 
discussed in sec. \ref{ssec:algorithm}. The order paratemer for the MAT2A system 
is defined to be $dis(OC)-dis(SC)$ and that for the $pv$ADA is defined as 
$dis(NC)-dis(OC)$. For the MAT2A system, configurations are sampled within 
the range $[-2.0,2.0]$ and this range is divided into 14 overlapping windows 
($0.09$ {\AA} overlap on either side). 
Configurations are sampled for the $pv$ADA system within the interval $[-3.0, 3.0]$ {\AA}
of the order parameter value which is divided into 20 overlapping 
windows ($0.045$ {\AA} between neighboring windows). A reactive 
trajectory from the TPS ensemble is chosen to be the guide trajectory 
for sampling within each of the windows. To sample in a particular window $w$, a 
frame from this reactive trajectory is chosen such that the value of its 
order parameter falls within $w$. Using the shooting algorithm
the momenta are perturbed and dynamics is run for 10 steps forwards and backwards with a 
$1\;fs$ time step to obtain a trajectory with a time period of $20\;fs$.    
Within each window 3000 TPS trajectories are collected. The 
probability distribution function for the order parameter within window $w$ 
is then calculated as normalized histograms which represent the populations 
of the configurations within the window at equilibrium. 
Then we calculate the free energy profile in a similar fashion to the WHAM 
procedure \cite{Kumar92JComputChem13p1011} which is generally used to obtain the 
potential of mean force from umbrella sampling calculations. 
%------------------------------------------------------------------------------ 
\section{Results and discussion}
%------------------------------------------------------------------------------ 
\subsection{Human Methionine Adenosyltransferase 2A}
The equilibrated structure of the MAT2A enzyme in complex with 
ATP, MET and two Mg$^{2+}$ ions is shown in Fig. \ref{fig:mat2a-equil}.
\begin{figure}[ht!]
\centering
\includegraphics[scale=0.2]{figures/mat2a-equil.png}
\caption{Equilibriated structure of the MAT2A enzyme in complex with ATP and 
MET along with 2 Mg$^{2+}$ ions.}
\label{fig:mat2a-equil}
\end{figure}
In the binding pocket the adenosyl moiety and the Phe250 and Ile117 residues 
are found to have $\pi-\pi$ interactions stabilizing the molecule. The triphosphate 
moiety forms hydrogen bonds with ... residues.

\begin{figure}
\includegraphics[scale=1.0]{figures/mat2a-diff167.pdf}
\caption{Bond breaking ($\mathrm{d}_{\mathrm{OC}}$), bond forming 
($\mathrm{d}_{\mathrm{SC}}$) distances for a typical 
reactive trajectory for the reaction catalyzed by the MAT2A enzyme.}
\label{fig:mat2a-reactive-traj}
\end{figure}
The order parameters of the system along a typical reactive trajectory is 
shown in Fig. \ref{fig:mat2a-reactive-traj}.

Committor analysis and committor distribution analysis are used to deduce 
the transition states and the reaction coordinates, respectively. 
Committor analysis is carried out for a reactive trajectory 
from the TPS ensemble by choosing a time slice randomly from the trajectory, assigning
momenta picked at random from the Boltzmann distribution for all the atoms, and running 
dynamics for $250\;fs$. The commitment probability of a time slice is calculated by 
repeating this procedure several times and finding the fraction of trajectories that go 
to the reactant, product or neither. The transition state for the reaction is defined as 
the structure from the time slice that has equal probability to commit to the reactant 
and the product state. Repeating this process for several trajectories from the TPS
ensemble results in a collection of transition states called the separatrix. 
The reaction coordinate might contain not just the atoms from the molecules that 
are directly involved in the reaction but also residues from side chains that can 
promote the reaction. Taking this into account the committor distribution analysis
starts by constraining the residues and molecules that are hypothesized to be part of the 
reaction coordinate and using this constrained system as the starting point 
trajectories are launched assigning momenta at random from the Boltzmann distribution at 
$300\;\text{K}$.

\begin{figure}[ht!]
\includegraphics[scale=0.12]{blender-images/ada/mat2a-trans-labelled.png}
\caption{Transition state structure of the reaction catalyzed by the human MAT2A enzyme.
The S$-$C distance is $2.09$ {\AA} and the O$-$C distance is $2.32$ {\AA} which represents 
an advanced transition state.}
\label{fig:mat2a-trans-struct}
\end{figure}
The transition state structure follows a S$_{\text{N}}2$ 
mechanism with a product like structure, the O$-$C bond is cleaved and 
has a value of $2.32$ {\AA}.
Experimentally the transition state structures of enzymatic reactions
can be elucidated using kinetic isotope effects (KIE). \cite{Schramm99MetEnzym308p301}

From the committor distribution analysis for a reactive trajectory we see that the 
constrained region consisting of just the QM part which is the ATP and MET molecules 
along with two Mg$^{2+}$ ions results in a distribution that is not peaked 
at 0.5 whereas when the residues ... are constrained we obtain a distribution 
that is peaked at 0.5. Therefore we conclude that the reaction coordinate consists
of these residues as well. 
\begin{figure*}[ht!]
\centering
\begin{minipage}[b]{0.45\linewidth}
\includegraphics[width=\textwidth]{figures/comm-60-mat2a-nocons.pdf}
\label{fig:minipage1}
\end{minipage}
\quad
\begin{minipage}[b]{0.45\linewidth}
\includegraphics[width=\textwidth]{figures/comm-60-mat2a.pdf}
\label{fig:minipage2}
\end{minipage}
\caption{Committor distribution analysis for obtaining the reaction coordinate of the 
MAT2A catalyzed reaction. The figure on the left has the QM region constrained and the figure 
on the right has the QM region along with the Gln113, Ser114, Arg249 and Arg264 residues constrained.}
\end{figure*}

@@

@@

From our TPS based calculations we calculated the free energy barrier to be 
$16\;\text{Kcal mol}^{-1}$.

@@

\text{Kcal mol}^{-1}$. Compared to typical enzyme catalyzed 
reactions calculated activation energy is small.  

@@

\subsection{Adenosine deaminase reaction coordinates and free energy}

@@

The bond forming ($dis(OC)$) and bond breaking ($dis(NC)$) distances 

@@

where intermediate Meisenheimer complex is formed at the transition state
as shown in Fig. \ref{fig:ada-trans}. 
The approaching O atom of the hydroxide anion is at a distance of $1.48$ {\AA}

@@

The formation of the Meissenheimer complex is the rate limiting step and the N-C distance is 
1.60 {\AA} and the O-C distance is 1.48 {\AA}.}

@@

catalyzed by the plasmodium falciparum ADA enzyme which shares
$72\%$ identity with the $pv$ADA enzyme that we study here. The frequency of the 
configurations (order parameter values) that are calculated within each of the 
windows is used to  

@@

The free energy barrier is calculated to be $23\;\text{Kcal}\;\text{mol}^{-1}$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  Transition path sampling (TPS) is widely used for the
  calculations of reaction rates, transition state structures and reaction 
  coordinates of condensed phase systems. Here we discuss a scheme 
  for the calculation of free energies using the ensemble of TPS reactive 
  trajectories in combination with a window based sampling technique 
  %introduced by 
  %Radhakrishnan et al. (Radhakrishnan, R.; Schlick, T., J. Chem. Phys. 2004, 121, 2436.) 
  for enzyme catalyzed reactions. We calculate the free energy profiles of 
  the reactions catalyzed by the human methionine S-adenosyltransferase (MAT2A) 
  enzyme and the plasmodium vivax adenosine deaminase enzyme to assess 
  the accuracy of this method. MAT2A catalyzes the formation of S-adenosine-L-methionine
  following a S$_{\text{N}}2$ mechanism and using our method we estimate
  the free energy barrier to be $16\;\text{Kcal}\;\text{mol}^{-1}$ which
  is in excellent agreement with the experimentally measured activation 
  energy of $17.27\;\text{Kcal}\;\text{mol}^{-1}$. Furthermore for the 
  $pv$ADA enzyme catalyzed reaction we estimate a free energy barrier of 
  $23\;\text{Kcal}\;\text{mol}^{-1}$ and the calcualted free energy profile is
  similar to that predicted from experimental observations.
  We show that calculating free energies within TPS is lucrative as it is 
  bias free, accurate compared to experimental measurements 
  and computationally reasonable.    
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Molecular dynamics (MD) simulations are increasingly being
used to calculate quantitative estimations of experimentally measurable 
kinetic and thermodynamic parameters of enzyme catalyzed reactions. 
\cite{Karplus02NatStructMolBiol9p646,
Schramm98AnnuRevBiochem67p693,Zhang05AccChemRes38p379,
Schramm11AnnuRevBiochem80p703,Schramm18ChemRev118p11194} 
This has helped in proposing mechanisms and explanations for 
enzyme activity, fast high throughput screening of thousands of inhibitor 
molecues in drug lead discovery, \cite{Jorgensen09AccChemRes42p724,Sliwoski14PharmacolRev66p334} 
directed evolution for designing enzymes with enhanced catalytic 
activities, \cite{Thyme09Nature461p1300,Bloom09PNAS106p9995,Schafer19JAmChemSoc141p10431} etc.    
Straightforward MD simulations adequately sample 
the long-lived stable reactant and product states 
but the high energy barrier-crossing events that occur rarely and at long 
time scales are not easily accessible using this method alone.
To access such states at a higher frequency using MD simulations 
several enhanced sampling techniques 
have been developed in the past few decades some of which are 
umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
steered MD, \cite{Park04JChemPhys120p5946}
milestoning, \cite{Faradjian04JChemPhys120p10880} and transition path sampling 
(TPS). \cite{Pratt86JChemPhys85p5045,Bolhuis02AnnRevPhysChem53p291,
Dellago98JChemPhys108p1964,Bolhuis21AdvTheory4p2000237}

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{figures/mat2a-reaction.png}
\caption{The reaction catalyzed by the human MAT2A enzyme between ATP and L-methionine resulting in the 
formation of S-adenosyl L-methionine and triphosphate leaving group.}
\label{fig:mat2a-reaction}
\end{figure}

$pv$ADA catalyzes the reaction between adenosine and OH$^{-}$ ion 
resulting in the formation of inosine and ammonia as shown in
Fig. \ref{fig:ada-reaction}.
The nucleophilic O atom of the OH$^{-}$ anion
attacks the electrophilic C6 atom of the adenosyl ring leading to the 
formation of a Meisenheimer complex like structure while the N1 atom 
on the adenosyl moiety gets protonated by the Glu229 residue of $pv$ADA. 
The key bond parameters are the distance between the approaching hydroxyl 
oxygen and the C6 atom (denoted as d$_{\text{OC}}$) and the 
dissociating C6$-$N6 (d$_{\text{NC}}$) bond distance. For the TPS simulaions 
we define the reactant state to have d$_{\text{NC}}$ being less than $1.7$ {\AA} 
and the product state to have the d$_{\text{OC}}$ less than $1.7$ {\AA}.

\begin{figure}
\includegraphics[scale=0.6]{figures/ada-new.png}
\caption{The reaction catalyzed by the $pv$ADA enzyme between adenosine and hydroxide ion
resulting in the formation of inosine and ammonia.}
\label{fig:ada-reaction}
\end{figure}

A biased initial reactive trajectory is obtained by using harmonic restraint forces 
for both the MAT2A and $pv$ADA systems. 
Using the biased trajectory as the starting point and after removing the restraint 
forces, the shooting algorithm \cite{dellago02AdvChemPhys123} was used to generate
new trajectories. The shooting algorithm begins with the choice of a 
random time slice from the biased initial reactive trajectory and perturbs 
the momenta of all the atoms in the system. Then the dynamics is propagated for 
250 steps forawards and backwards using $1\;fs$ time steps
resulting in a trajectory with a time period of $0.5\;ps$.
A total of 150 reactive trajectories were generated for both the MAT2A and $pv$ADA
systems.
\subsection{Free energy calculations}
We follow the algorithm discussed previously to calculate the free energy
profile for the two enzymatic reactions using the TPS reactive trajectories. 
The order paratemer for the MAT2A system 
is defined to be d$_{\text{OC}}-$d$_{\text{SC}}$ and that for the $pv$ADA is defined as 
d$_{\text{NC}}-$d$_{\text{OC}}$. For the MAT2A system, configurations are sampled
for order parameter values within the range $[-2.0,2.0]$ and this range is divided 
into 14 overlapping windows ($0.09$ {\AA} between neighboring windows). 
Configurations are sampled for the $pv$ADA system for order parameter values 
within the interval $[-3.0, 3.0]$ {\AA} which is divided into 20 overlapping 
windows ($0.09$ {\AA} between neighboring windows). A reactive 
trajectory from the TPS ensemble is chosen to guide the sampling 
within each of the windows in the same spirit as a steered MD trajectory 
is used to guide the window based umbrella sampling simulations. 
To sample in a particular window $w$, a frame from this reactive 
trajectory is chosen such that the value of its 
order parameter falls within $w$. Using the shooting algorithm
the momenta of all the atoms in the system are perturbed 
and dynamics is run for 10 steps forwards and backwards with a 
$1\;fs$ time step to obtain a trajectory with a time period of $20\;fs$.
Longer trajectories in our experience end up sampling outside the window 
more often and hence we make use of very short trajectories in this study.     
Within each window 3000 TPS trajectories are collected. The 
probability distribution function for the order parameter within window $w$ 
is then calculated as normalized histograms which represent the populations 
of the configurations within the window at equilibrium. 
Finally we calculate the free energy profile in a similar fashion to the WHAM 
procedure \cite{Kumar92JComputChem13p1011} which is generally used to obtain the 
potential of mean force from umbrella sampling calculations. 
\section{Results and discussion}
\subsection{Human MAT2A}
The equilibrated structure of the MAT2A enzyme in complex with 
ATP, MET and two Mg$^{2+}$ ions is shown in Fig. \ref{fig:mat2a-equil}.
\includegraphics[scale=0.2]{figures/mat2a-equil.png}
\caption{Equilibriated structure of the MAT2A enzyme in complex with ATP and 
MET along with 2 Mg$^{2+}$ ions.}
\label{fig:mat2a-equil}
The variation of the key distance parameters of the reaction calculated for 
a reactive trajectory from the TPS ensemble are shown in Fig. \ref{fig:mat2a-reactive-traj}.

@@

\includegraphics[scale=1.0]{figures/mat2a-diff167.pdf}
\caption{Bond breaking ($\mathrm{d}_{\mathrm{OC}}$), bond forming 
($\mathrm{d}_{\mathrm{SC}}$) distances for a typical 
reactive trajectory for the reaction catalyzed by the MAT2A enzyme.}
\label{fig:mat2a-reactive-traj}

@@

Committor analysis and committor distribution analysis are used to deduce 
the transition states and the reaction coordinates, respectively. 
Committor analysis is carried out for a reactive trajectory 
from the TPS ensemble by choosing a time slice randomly from the trajectory, assigning
momenta picked at random from the Boltzmann distribution for all the atoms, and running 
dynamics for $250\;fs$. The commitment probability of a time slice is calculated by 
repeating this procedure several times and finding the fraction of trajectories that go 
to the reactant, product or neither. The transition state for the reaction is defined as 
the structure from the time slice that has equal probability to commit to the reactant 
and the product state. Repeating this process for several trajectories from the TPS
ensemble results in a collection of transition states called the separatrix.
\includegraphics[scale=0.12]{blender-images/ada/mat2a-trans-labelled.png}
\caption{Transition state structure of the reaction catalyzed by the human MAT2A enzyme.
The bond parameters d$_{\text{SC}}$ and d$_{\text{OC}}$ are $2.31$ {\AA} and $2.09$ {\AA}, 
respectively.}
\label{fig:mat2a-trans-struct}
The transition state resembles a S$_{\text{N}}2$ structure with the 
d$_{\text{SC}}$ and d$_{\text{OC}}$ distances being $2.31$ {\AA} and $2.09$ {\AA}, respectively.
Experimentally the transition state structures of enzymatic reactions
can be elucidated using kinetic isotope effects (KIE). \cite{Schramm99MetEnzym308p301}
Firestone et al. \cite{Firestone17JAmChemSoc139p13754} used a combination of experimental 
KIE measurements using the whole protein system and computational density functional 
theory (DFT) based gas phase simulations of representative molecules to elucidate the 
transition state structure for this reaction. From their findings they report 
d$_{\text{SC}}$ and d$_{\text{OC}}$ distances of $2.03$ {\AA} and $2.32$ {\AA}, 
respectively and our calculated values are in close agreement to this result.

The reaction coordinate might contain not just the atoms from the molecules that 
are directly involved in the reaction but also residues from side chains that can 
promote the reaction. Taking this into account the committor distribution analysis
starts by constraining the residues and molecules that are hypothesized to be part of the 
reaction coordinate and using this constrained system as the starting point 
trajectories are launched assigning momenta at random from the Boltzmann distribution at 
$300\;\text{K}$ to all the atoms in the system. 
First we constrained just the QM region in our committor distribution analysis 
which consists of ATP and MET molecules along with two Mg$^{2+}$ ions and the 
resulting distribution as shown in Fig. \ref{fig:mat2a-comm-dist} is not peaked 
at 0.5 which is expected for the reaction coordinate. 
Next we constrained the residues Gln113, Ser114, Arg249 and Arg264 along with the 
QM region and obtain a distribution that is peaked at 0.5 and hence we 
conclude that the reaction coordinate consists of these residues as well. 
\begin{figure*}[ht!]
\begin{minipage}[b]{0.45\linewidth}
\includegraphics[width=\textwidth]{figures/comm-60-mat2a-nocons.pdf}
\label{fig:minipage1}
\end{minipage}
\quad
\begin{minipage}[b]{0.45\linewidth}
\includegraphics[width=\textwidth]{figures/comm-60-mat2a.pdf}
\label{fig:minipage2}
\end{minipage}
\caption{Committor distribution analysis for obtaining the reaction coordinate of the 
MAT2A catalyzed reaction. The figure on the left has the QM region constrained and the figure 
on the right has the QM region along with the Gln113, Ser114, Arg249 and Arg264 residues constrained.}
\label{fig:mat2a-comm-dist}
\end{figure*}

@@

\text{Kcal mol}^{-1}$ which is small Compared to typical enzyme catalyzed 
reactions.
\label{fig:mat2a-fenergy}
The free energy profile from our TPS based calculations is shown in 
Fig. \ref{fig:mat2a-fenergy} where the barrier is $16\;\text{Kcal mol}^{-1}$.
This is in excellent agreement with the experimental result.  

@@

\subsection{$pv$ADA}

@@

The bond forming (d$_{\text{OC}}$) and bond breaking (d$_{\text{NC}}$) distances 

@@

where an intermediate Meisenheimer complex like structure is formed 
at the transition state as shown in Fig. \ref{fig:ada-trans}. 
The approaching O atom of the OH$^{-}$ anion is at a distance of $1.48$ {\AA}

@@

The formation of the Meissenheimer complex like structure with 
d$_{\text{NC}}=1.60$ {\AA} and d$_{\text{OC}}=1.48$ {\AA} is observed.}

@@

catalyzed by the plasmodium falciparum ADA ($pf$ADA) enzyme which shares
$72\%$ identity with the $pv$ADA enzyme that we analyze here. The next
step in the reaction is the proton transfer from the OH$^{-}$ ion to the 
amine leaving group resulting in the formation of inosine and ammonia.   
%At the transition state, experimental findings 
%suggest that the dissociation of the C6$-$N6 bond as well as the protonation of the N1 
%atom are almost complete. \cite{Luo07JAmChemSoc129p8008}

@@

The free energy barrier is calculated to be $23\;\text{Kcal}\;\text{mol}^{-1}$
and the overall shape of the function is very similar to that hypothesized 
by Luo et al. \cite{Luo07JAmChemSoc129p8008}
